Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2016-06-01 draft.tex
misc changes from editor, including fig. 2

\documentclass[11pt,letterpaper, notitlepage]{article}
\usepackage{epsf,graphicx,psfrag}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage[reqno]{amsmath}
\usepackage{endnotes}
\usepackage{natbib}
\usepackage{amsfonts}
\usepackage{amssymb,enumerate}
\usepackage{dcolumn}
\usepackage{verbatim}
%\usepackage{hypertex}
\usepackage[pdftex, bookmarks, colorlinks=true, plainpages=false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[hmargin=3.0cm,vmargin=3.0cm]{geometry}
%\usepackage{geometry}
\usepackage{nicefrac}
\usepackage{pdfsync}
\usepackage{url}
\usepackage{pst-all}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage{graphicx,verbatim,scalefnt}
\pdfoptionpdfminorversion=6 %gets rid of warning about pdf version

@@

The payoffs at the end of a period are determined by these actions and six state variables, which keep track of the per-period value of the status quo ($Q_s^t$ and $Q_g^t$), losing the issue of recognition ($L_s^t$ and $L_g^t$) and winning the issue of recognition ($W_s^t$ and $W_g^t$) to the secessionists and government respectively. The game continues into the next period if the inside players selected (status quo, status quo) or (cede, cede). Otherwise, the game ends in an absorbing state as described below.

@@

The other payoffs for the inside actors are also modeled as state variables. However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g. $L_s^t = L_s^{t-1} - R_p^t + R_c^t$ where $L_s^t$ is the per-period payoff to the secessionists from ceding the issue of status. The full set of payoffs for the inside actors are described in Section~\ref{sec:stage}.

Both because of the presence of absorbing states and these state variables whose values depend on their values in the previous period, this is a dynamic game rather than a standard repeated game.

Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates. This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} U_i^t \de^{t-1}$.

In each period, play proceeds as follows:

@@

\item $p$ chooses a level of resource expenditure $R_p^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.

\item $c$ chooses a level of resource expenditure $R_c^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.

@@

\noindent Stage game payoffs for players $s$ and $g$ in period $n$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $\Omega^t$ &$\Omega^t$&$W_g^t, L_s^t$ \\ \hline
	Status Quo& $\Omega^t$& $Q_g^t, Q_s^t$ &$W_g^t, L_s^t$ \\ \hline
	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$Q_g^t, Q_s^t$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
\vspace{2.5pt}

@@

If both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period. In this case neither player has demonstrated a willingness to give up more than the other. Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}

@@

If both $s$ and $g$ play Status Quo, then the status quo persists.

@@

At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game. For example, if $p$ expends resources $R_p^t$ to increase the government's payoffs from ceding, the previous period's payoff $L_g^{t-1}$ is augmented so that $L_g^t = L_g^{t-1} + R_p^t$. We can write analogous expressions for any of the state variables, recalling that the international community can also make investments to increase whichever of these payoffs they like. $B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.

>>>>

\documentclass[11pt,letterpaper, notitlepage]{article}
\usepackage{epsf,graphicx,psfrag}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage[reqno]{amsmath}
\usepackage{endnotes}
\usepackage{natbib}
\usepackage{amsfonts}
\usepackage{amssymb,enumerate}
\usepackage{dcolumn}
\usepackage{verbatim}
%\usepackage{hypertex}
\usepackage[pdftex, bookmarks, colorlinks=true, plainpages=false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[hmargin=3.0cm,vmargin=3.0cm]{geometry}
%\usepackage{geometry}
\usepackage{nicefrac}
\usepackage{pdfsync}
\usepackage{url,bm}
\usepackage{pst-all}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage{graphicx,verbatim,scalefnt}
\pdfoptionpdfminorversion=6 %gets rid of warning about pdf version

@@

The payoffs at the end of a period are determined by these actions and six state variables, which keep track of the evolving value of the status quo ($Q_s^t$ and $Q_g^t$), losing the issue of recognition ($L_s^t$ and $L_g^t$) and winning the issue of recognition ($W_s^t$ and $W_g^t$) to the secessionists and government respectively. The game continues into the next period if the inside players selected (status quo, status quo) or (cede, cede). Otherwise, the game ends in an absorbing state as described below.

@@

The other payoffs for the inside actors are also modeled as state variables. However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, to be precise $L_s^t = L_s^{t-1} - R_p^t + R_c^t$ where $L_s^t$ is the per-period payoff to the secessionists from ceding the issue of status. The full set of payoffs for the inside actors are described in Section~\ref{sec:stage}.

Both because of the presence of absorbing states and state variables, this is a dynamic game rather than a standard repeated game.

Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates. This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.

In each period $t$, play proceeds as follows:

@@

\item $p$ chooses an investment level $R_p^t\in \mathbb{R}^+$.

\item $c$ chooses an investment level $R_c^t\in \mathbb{R}^+$.

@@

\noindent Stage game payoffs for players $s$ and $g$ in period $n$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
\vspace{2.5pt}

@@

@@

Only two outcomes of the stage game do not lead to absorbing states, and the payoffs for these are bolded. If both $s$ and $g$ play Status Quo, then the status quo persists. Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period. In this case neither player has demonstrated a willingness to give up more than the other. Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}

@@

At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game. For example, if $p$ expends resources $R_p^t$ to increase the government's payoffs from ceding, the previous period's payoff $L_g^{t-1}$ is augmented so that $L_g^t = L_g^{t-1} + R_p^t - R_c^t$. We can write analogous expressions for any of the state variables, recalling that the international community can also make investments to increase whichever of these payoffs they like. $B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
